Redundant calibration as a complex optimization problem

### 2016-12-14 paper.tex
added new text to intro

\section{Introduction}
The propogation effects (i.e. the ionosphere, antenna gains, etc.) a celestial signal experiences along its propogation path is modelled with the radio interforemetric measurement equation (RIME). %\citep{ME1,RRIME1}.
Calibration is the procedure by which we remove the errors that these propogation effects induce onto our intereferometric data and is normally accomplished using a 
least-squares solver. In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimizes the difference betweem our observed and predicted visibilities. There are two recent developments
in calibration which should be explicitly mentioned, since they form the foundation this paper builds on. The first is the discovery of StEfCal \citep{salvini2014}, which 
is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates. Now we do the reverse,
assume the conugates are known and solve for the gains. This procedure results in the linearization of the calibration algorithm and reduces the computational complexity 
of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^3)$, where $N$ denotes the number of antennas in the array. The second is to not break the calibration problem 
into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters seperately, but to rather treat the antenna gains and their 
conjugates as seperate parameters to estimate \citep{smirnov2015}. \citet{smirnov2015} also showed that the StEfCal algorithm itself is merely a special case of this
gain-and-conjugate framework.

Redundant calibration differs from normal calibration in that it does not require a skymodel. With redundant calibration we do not only solve for the antenna gains, but also 
for the true observed visibilities. We are able to do so, because we have less unkonws than baseline equations. Since we

\subsection{Historical Overview}

@@

Main contributions of the paper:

\begin{enumerate}
 \item Present a general framework which unifies all the techniques developed so far showing they are all related (LINCAL, non-linear estimator, etc ...). They are all non-linear
 least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with LINCAL showing that it is GN, then we have to motivate 
 why we want to use Oleg's complex optimization framework.
 \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
 framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for LINCAL). State that the algorithm is 
 effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
 \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
 \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
 conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
 we explore both of these parameters.
 \end{enumerate}

 **************
 FLOW OF PAPER
 **************

 \begin{enumerate}
 \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
 \item Introduce the array geometery function.
 \item Write down logcal and lincal in matrix form... ?
 \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
 \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
 \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
 parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
 \item Show how this relates to lincal for example ---> show lincal is GN.
 \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
 we can derive the method of Chengular.
 \item Do the derivation of Chengular. 
 \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
 on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
 idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
 approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
 relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
 to how Oleg re-derived stefcal in, we have achieved the same approach.
 \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
 implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
 full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
 gradient method.
 \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
 are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
 cg. 
 \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
 \item Simulation description
 \item Will CG improve things?
 \item Now first discuss the condition
 number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
 the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
 \item Now we discuss the sparsity results. 
 \item Provide a table that theoretical compares R-StEFCal and SPARC.
 \item Number of outerloop iterations. 
 \item Timing results.
 \item Accuracy results.
 \item Maybe some freq simulations.
 \end{enumerate}

 We

>>>>

\section{Introduction}
The propogation effects (i.e. the ionosphere, antenna gains, etc.) a celestial signal experiences along its propogation path is modelled with the radio interforemetric measurement equation (RIME). %\citep{ME1,RRIME1}.
Calibration is the procedure by which we remove the errors that these propogation effects induce onto our intereferometric data and is normally accomplished using a 
least-squares solver. In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimizes the difference betweem our observed and predicted visibilities. There are two recent developments
in calibration which should be explicitly mentioned, since they form the foundation this paper builds on. The first is the discovery of StEfCal \citep{salvini2014}, which 
is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates. Now we do the reverse,
assume the conugates are known and solve for the gains. This procedure results in the linearization of the calibration algorithm and reduces the computational complexity 
of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^3)$, where $N$ denotes the number of antennas in the array. The second development is the idea; to not break the calibration problem 
into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters seperately, but to rather treat the antenna gains and their 
conjugates as seperate parameters to estimate \citep{smirnov2015}. \citet{smirnov2015} also showed that the StEfCal algorithm itself is merely a special case of this
gain-and-conjugate approach.

Two baselines are redundant if their beaseline difference vector have the same lenght and orientation, which imply that they measure the same $uv$-modes. When we use a redundant array layout we dramatically 
reduce the number of unkowns. Redundant calibration is the procedure by which we leverage the fact that we have 
Redundant calibration differs from normal calibration in that it does not require a skymodel. With redundant calibration we do not only solve for the antenna gains, but also 
for the true observed visibilities. We are able to do so, because we have less unkonws than baseline equations. Since we

@@

EOR SECTION SOMEWHERE HERE

MAIN CONTRIBUTION OF THIS PAPER COMES HERE

% Main contributions of the paper:
% 
% \begin{enumerate}
%  \item Present a general framework which unifies all the techniques developed so far showing they are all related (LINCAL, non-linear estimator, etc ...). They are all non-linear
%  least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with LINCAL showing that it is GN, then we have to motivate 
%  why we want to use Oleg's complex optimization framework.
%  \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
%  framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for LINCAL). State that the algorithm is 
%  effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
%  \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
%  \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
%  conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
%  we explore both of these parameters.
%  \end{enumerate}
%  
%  **************
%  FLOW OF PAPER
%  **************
%  
%  \begin{enumerate}
%  \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
%  \item Introduce the array geometery function.
%  \item Write down logcal and lincal in matrix form... ?
%  \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
%  \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
%  \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
%  parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
%  \item Show how this relates to lincal for example ---> show lincal is GN.
%  \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
%  we can derive the method of Chengular.
%  \item Do the derivation of Chengular. 
%  \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
%  on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
%  idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
%  approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
%  relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
%  to how Oleg re-derived stefcal in, we have achieved the same approach.
%  \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
%  implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
%  full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
%  gradient method.
%  \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
%  are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
%  cg. 
%  \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
%  \item Simulation description
%  \item Will CG improve things?
%  \item Now first discuss the condition
%  number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
%  the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
%  \item Now we discuss the sparsity results. 
%  \item Provide a table that theoretical compares R-StEFCal and SPARC.
%  \item Number of outerloop iterations. 
%  \item Timing results.
%  \item Accuracy results.
%  \item Maybe some freq simulations.
%  \end{enumerate}
%  
%  We
